---
title: "Do weather affect driving patterns?"
author: "Sindre Sønvisen"
date: "10/18/2020"
output: 
  html_document:
    df_print: paged
---

```{css, comment=NA, message=FALSE, echo=FALSE, warning=FALSE}
body {background-color: lightgray;}
```

```{r setup, include=FALSE, comment=NA, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("wrangle_new.r")
source("regression.r")
library(leaflet) #Map
```

# Introduction:
<pre>
  This project is about testing whether the weather has a say in the  
  number of available parking spaces in Stavanger.
</pre>

# Data:
<pre>
  The weather data is collected from:
    Norwegian klima service senter:
      https://klimaservicesenter.no/observations/
    Temperature and Precipitation from station id:SN44640.
    Wind information from station id:SN44560. (consider changing...)
  The parking data is collected from:
    https://open.stavanger.kommune.no/dataset/stavanger-parkering/
    
  More information on each parking lot:
    https://stavanger-parkering.no/parkering/p-hus/
</pre>

# Hypothesis:
<pre>
  The available parking spaces goes down in bad weather.
</pre>

## Map
```{r map, comment=NA, message=FALSE, echo=FALSE, warning=FALSE}
map3 <- leaflet() %>%
addTiles() #%>%  # Add default OpenStreetMap map tiles
for(n in 1:n_places){
  temp <- loc_parking[n,]
  map3 <- map3 %>% addMarkers(lat=temp$Lat, lng=temp$Lon, popup=temp$Place)
}
map3
```

## Correlation
A schema of correlation between the parking lots and the different weather aspects.
```{r Correlation, comment=NA, message=FALSE, echo=FALSE, warning=FALSE}
df_temp <- df_main %>% 
  filter(hour(DateTime) >= 8) %>%
  filter(hour(DateTime) <= 18) %>% .[-c(1)]
grid.table(
    cor(
      x=df_temp[,1:9], 
      y=df_temp[,c("Air_temp degC", "Precipitation mm", "wind_speed m/s")], 
      use = "complete.obs", 
      method = c("pearson")
      )
)
```

## Linear regression

```{r Regression, comment=NA, message=FALSE, echo=FALSE, warning=FALSE}
make_regression(df_main, "Air_temp degC")
make_regression(df_main, "Precipitation mm")
make_regression(df_main, "wind_speed m/s")
```

## Time graphs
```{r Time_p, comment=NA, message=FALSE, echo=FALSE, warning=FALSE}
df_parking %>% ungroup() %>%
  plot_ly(x = ~DateTime, y = ~Free_spaces, color = ~Place, mode = "lines") %>%
  layout(
    title = list(text = "Parking time graph", x = 0.1),
    yaxis = list(title = "Available parking spaces"),
    xaxis = list(Titel = "Date",
                 rangeselector = list(x = 0.01, y = 0.97,
                   buttons = list(
                     list(
                       count = 1,
                       label = "1 day",
                       step = "day",
                       stepmode = "backward"),
                     list(
                       count = 7,
                       label = "1 week",
                       step = "day",
                       stepmode = "backward"),
                     list(
                       count = 1,
                       label = "1 month",
                       step = "month",
                       stepmode = "backward"),
                     list(
                       count = 1,
                       label = "YTD",
                       step = "year",
                       stepmode = "todate"),
                     list(step = "all"))),
                 rangeslider = list(type = "date"))
    )

```

```{r Time_W, comment=NA, message=FALSE, echo=FALSE, warning=FALSE}
df_weather %>% filter(Element != "wind_direction degrees") %>% ungroup() %>%
  plot_ly(x = ~DateTime, y = ~Value, color = ~Element, mode = "lines") %>%
  layout(
    title = list(text = "weather time graph", x = 0.1),
    yaxis = list(title = "Value"),
    xaxis = list(Titel = "Date",
                 rangeselector = list(x = 0.01, y = 0.97, 
                   buttons = list(
                     list(
                       count = 1,
                       label = "1 day",
                       step = "day",
                       stepmode = "backward"),
                     list(
                       count = 7,
                       label = "1 week",
                       step = "day",
                       stepmode = "backward"),
                     list(
                       count = 1,
                       label = "1 month",
                       step = "month",
                       stepmode = "backward"),
                     list(
                       count = 1,
                       label = "YTD",
                       step = "year",
                       stepmode = "todate"),
                     list(step = "all"))),
                 rangeslider = list(type = "date"))
    )

```

# Methods used:
....